---
title: "Tidy Tuesday"
date: "week 29, 2021"
output:
  html_document:  
    df_print: paged
    css: styles.css
    highlight: tango
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
    number_sections: yes
---

```{r include = FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE,
	class.output="style-output",
  class.source="style-source"
)

options(
  digits = 1,
  rows.print = 20,
  cols.print = 4,
  paged.print = TRUE
)
```

The dataset this week comes from [Kaggle](https://www.kaggle.com/williamschooleman/scoobydoo-complete) and contains all Scooby-Doo episodes. More info about Scooby Doo can be found on [ScoobyPedia](https://scoobydoo.fandom.com/wiki/Scoobypedia).

# Setup

```{r}
pacman::p_load(
  tidytuesdayR,
  tidyverse,
  lubridate,
  magrittr,
  glue,
  skimr,
  here
)
```

# Raw Data

Get data and write to local file

```{r eval = FALSE}
tt_load(2021, week = 29) %>% 
  pluck("scoobydoo") %>% 
  write_csv2(glue("{here()}/data.csv"))
```

Read data from local file (`d_raw`)

```{r}
d_raw <- read_csv2(
  glue("{here()}/data.csv"),
  col_types = cols(.default = "c"),
  na = c("NA", "NULL")
)
```

Create working copy (`d`) and sort columns alphabetically

```{r}
d <- d_raw %>% select(sort(colnames(.)))
```

# Inspection

```{r}
d
```

```{r}
glimpse(d)
```

## Coltypes

Names of columns that appear logical at first glance

```{r}
# specify columns
col_lgl <- d %>% 
  select(c(
    "arrested",
    "batman",
    "blue_falcon",
    "door_gag",
    "hex_girls",
    "monster_real",
    "non_suspect",
    "scooby_dum",
    "scrappy_doo",
    "trap_work_first",
    starts_with("captured_"),
    starts_with("caught_"),
    starts_with("snack_"),
    starts_with("unmask_")
  )) %>% 
  colnames()

# inspect sample
d %>% select(col_lgl) %>% slice_sample(n = 15)
```

Names of columns that appear numeric at first glance

```{r}
# specify columns
col_num <- c(
  "another_mystery",
  "culprit_amount",
  "engagement",
  "groovy",
  "imdb",
  "index",
  "jeepers",
  "jinkies",
  "just_about_wrapped_up",
  "monster_amount",
  "my_glasses",
  "number_of_snacks",
  "rooby_rooby_roo",
  "run_time",
  "scooby_doo_where_are_you",
  "season",
  "set_a_trap",
  "split_up",
  "suspects_amount",
  "zoinks")

# inspect sample
d %>% select(col_num) %>% slice_sample(n = 15)
```

# Wrangling

## Coercion

### Date

Parse date from character vector `date_aired`

```{r}
# parse date
d %<>% mutate(aired_date = ymd(date_aired))

# verify result
d %>% 
  select(date_aired, aired_date) %>% 
  slice_sample(n = 15)

# drop original column
d %<>% select(-date_aired)
```

Extract date elements from `aired_date`

```{r}
# get year and month
d %<>% mutate(
  aired_year = year(aired_date),
  aired_month = month(aired_date, label = TRUE)
) 

# get decade
d %<>% mutate(
  aired_decade = glue("{floor(aired_year/10) * 10}s")
) %>% as_factor()

# verify result
d %>% select(
    aired_date,
    aired_year,
    aired_month,
    aired_decade) %>% 
  slice_sample(n = 15)
```

### Logical

Show all unique values in the columns that are presumed to be logical to manually verify this assumption.

```{r}
# print unique values
d %>% 
  select(col_lgl) %>% 
  map(~unique(.x)) %>% 
  flatten_chr() %>% 
  unique()
```

There are indeed no other values than `"TRUE"`/`"FALSE"` or `NA`.

```{r}
# convert to logical
d %<>% mutate(across(col_lgl, as.logical))

# verify result
d %>% select(col_lgl) %>% slice_sample(n = 15)
```

### Numeric

Do the (presumed) numeric columns contain non-numeric characters?

```{r}
# store unique values per column
uniq <- d %>% 
  select(col_num) %>% 
  map(~unique(.x)) 

# identify columns with non-numeric characters
non_num <- uniq %>%
  map( ~ str_detect(.x, "\\D") %>%
         replace_na(FALSE) %>%
         any()) %>%
  flatten_lgl()

# print unique values in columns with non-numeric characters
uniq[non_num]
```

The only non-numeric character in `imdb` is the decimal `.`, so this column can easily be converted to numeric. On the other hand, for `season` and `number_of_snacks` this doesn't make sense.

```{r}
# exclude season and number of snacks
col_num_final <- col_num[!(col_num %in% c("season", "number_of_snacks"))]
```

I like to keep the original columns (temporarily) for easy verification.

```{r}
# convert to numeric and add suffix
d_num <- d %>% 
  select(all_of(col_num_final)) %>%  
  map_dfr(as.numeric) %>% 
  rename_with( ~ glue("{.x}_num"))

# bind converted data to original data
d %<>% bind_cols(d_num)

# sort columns alphabetically
d %<>% select(sort(colnames(.)))

# verify result
d %>% select(sort(c(
  col_num_final, glue("{col_num_final}_num")
))) %>% slice_sample(n = 15)

# drop old columns and remove suffix from new columns
d %<>%
  select(-(col_num_final)) %>%
  rename_with(~ str_remove(.x, "_num"))
```

## Nested

Identify character columns containing ≥1 comma (might indicate nested data)

```{r}
d_chr <- d %>% select(where(is.character))
d_chr[d_chr %>%
        map( ~ any(str_detect(.x, ","))) %>%
        flatten_lgl() %>%
        replace_na(FALSE)]
```

Create nested list columns where applicable

```{r eval = FALSE}
# note to self: not working, figure out why...
col_nested <-
  c(
    "culprit_gender",
    "culprit_name",
    "monster_gender",
    "monster_name",
    "monster_species",
    "monster_type",
    "monster_subtype"
  )

d %<>% mutate(
  across(
    col_nested, 
    ~ str_squish(str_split(.x, ","))
  )
)
```

```{r}
# the non-elegant -but working- alternative
d$culprit_gender %<>% str_split(",")
d$culprit_name %<>% str_split(",")
d$monster_gender %<>% str_split(",")
d$monster_name %<>% str_split(",")
d$monster_species %<>% str_split(",")
d$monster_type %<>% str_split(",")
d$monster_subtype %<>% str_split(",")
```

# Clean Data

Result of the data cleaning and wrangling.

```{r}
d
```

# Exploration

## Skim

```{r}
skim(d)
```

# Visualization

## Monsters are male

```{r}
plotdata <- d %>%
  select(aired_decade, monster_gender) %>%
  unnest(monster_gender) %>%
  group_by(aired_decade, monster_gender) %>%
  filter(monster_gender %in% c("Male", "Female")) %>%
  tally()

plotdata %>% slice_sample(n = 10)

ggplot(plotdata, aes(
  fill = as.factor(monster_gender),
  y = n,
  x = as.factor(aired_decade)
)) +
  geom_bar(position = "fill", stat = "identity")
```

> ## ... and so are culprits!

```{r}
gd <- d %>% 
  select(aired_year, culprit_gender) %>% 
  unnest(culprit_gender) %>% 
  mutate(culprit_gender = str_squish(culprit_gender)) %>% 
  group_by(aired_year, culprit_gender) %>% 
  filter(!(culprit_gender %in% c("", "None"))) %>% 
  tally()

ggplot(gd, aes(
  fill = as.factor(culprit_gender),
  y = n,
  x = as.factor(aired_year)
)) + geom_bar(position = "fill", stat = "identity")
```

## 

```{r}
gd <- d %>%
  select(starts_with("unmask_")) %>%
  rename_with( ~ str_remove(.x, "unmask_")) %>%
  pivot_longer(1:6, names_to = "name") %>% 
  group_by(name, value) %>% 
  filter(!is.na(value)) %>% 
  tally()

ggplot(gd, aes(x="", y = n, fill = value)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start = 0) + 
  facet_wrap(~ name, ncol = 2) +
  theme_void()
```
